Dewlap color variation in \textit{Anolis sagrei} is maintained among habitats within islands of the West Indies

### 2021-07-26 ms/sections/appendix.tex
v3.2

% Little but detectable differences on five islands

We tested for variation in \textit{A. sagrei} dewlap coloration between populations living in three characteristic habitat-types across nine islands that span the West Indian range of the brown anole (beach scrub, primary coppice and mangroves). We found that most of the variation in coloration was partitioned between islands (two-way PERMANOVA, $F(df = 8) = 43.7$, $P = 0.0001$, explained variance $R^2 = 40.9$\%). Nonetheless, we did find evidence for differences in dewlap coloration between habitat-types, and those were mostly island-specific (habitat-by-island interaction term, $F(16) = 3.53$, $P = 0.0001$, $R^2 = 6.6$\%), with a significant portion of the variation explained by an habitat effect across all islands, but this effect was relatively small ($F(2) = 4.7$, $P = 0.0001$, $R^2 = 1.1$\%).\\

We subsequently tested for differences in dewlap coloration between habitat-populations within each island, using within-island principal component scores (to maximize the variation captured for each island, see Methods). Our within-island random forest classification analyses revealed detectable differences in dewlap coloration on eight out of the nine islands in our sample: Abaco, Bimini, Cayman Brac, Eleuthera, Little Cayman, Long Island, North Andros and South Andros. The accuracy of random forest classification exceeded random expectation more often than expected by chance for all these islands (Table \ref{tab:randomforests}). Accuracy was as high as 73\% for Cayman Brac. We obtained similar results using other machine learning approaches such as support vector machines (Table \ref{tab:ksvms}) and linear discriminant analysis (Table \ref{tab:ldas}), except that these methods did not detect significant differences on Eleuthera and North Andros. We did not find evidence of spatial autocorrelation in dewlap coloration between the sites within islands, except for Eleuthera (Table \ref{tab:autocorrelation}). We now describe the specific differences detected on each island.\\

On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the beach scrub habitat were often mistaken for dewlaps from the coppice habitat (Fig. \ref{fig:Abaco}D). Importance analysis revealed that beach scrub and mangrove lizards mostly differed in reflectance in the ultraviolet (UV) end of the spectrum (below 400nm, Fig. \ref{fig:Abaco_supplement}F), where mangrove dewlaps had higher UV reflectance relative to beach scrub lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig. \ref{fig:Abaco}B). Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats (Fig. \ref{fig:Abaco}E, Table \ref{tab:anova}), representing a higher UV-reflectance relative to red (Fig. \ref{fig:Abaco}C). Beach scrub lizards also scored lower on PC3 (Fig. \ref{fig:Abaco}E, Table \ref{tab:anova}), indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig. \ref{fig:Abaco}C). Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig. \ref{fig:Abaco_supplement}G).\\

@@

Eleuthera was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors. Although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small. Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig. \ref{fig:Eleuthera}C, E, Table \ref{tab:anova}).\\

>>>>

Here we describe more precisely the patterns identified on each island.\\

On Abaco, dewlaps from the mangrove habitat were the best discriminated, while dewlaps from the beach scrub habitat were often mistaken for dewlaps from the coppice habitat (Fig. \ref{fig:Abaco}D). Importance analysis revealed that beach scrub and mangrove lizards mostly differed in reflectance in the ultraviolet (UV) end of the spectrum (below 400nm, Fig. \ref{fig:Abaco_supplement}F), where mangrove dewlaps had higher UV reflectance relative to beach scrub lizards, and coppice lizards had an intermediate UV reflectance between the two other habitats (Fig. \ref{fig:Abaco}B). Consistent with this, our analyses of variance detected significantly lower PC2 scores in mangrove lizards than in the two other habitats (Fig. \ref{fig:Abaco}E, Table \ref{tab:anova}), representing a higher UV-reflectance relative to red (Fig. \ref{fig:Abaco}C). Beach scrub lizards also scored lower on PC3 (Fig. \ref{fig:Abaco}E, Table \ref{tab:anova}), indicating less curvature of the reflectance profile and relatively higher reflectance at intermediate wavelengths (blue-to-yellow) than at the ends of the range (Fig. \ref{fig:Abaco}C). Differences were detected between sites both at large ($\sim$ 100km) and short ($<$ 1km) distances (Fig. \ref{fig:Abaco_supplement}G). Abaco was the only island where we detected significant spatial autocorrelation (Table \ref{tab:autocorrelation}), that is, sites that were closer geographically tended to have populations of lizards with more similar dewlap colors.\\

@@

On Eleuthera, although random forests detected between-habitat differences in dewlap color, other approaches did not (Tables \ref{tab:ldas} and \ref{tab:ksvms}), suggesting that the differences may be small. Consistent with this, the only significant univariate difference detected was for PC2 between beach scrub and mangrove lizards, where beach scrub lizards had higher levels of red reflectance and mangrove lizards higher levels of UV reflectance (Fig. \ref{fig:Eleuthera}C, E, Table \ref{tab:anova}).\\

### 2021-07-26 ms/sections/discussion2.tex
v3.2

Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers). Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpubl.), and (3) \textit{A. sagrei} has been shown to be a highly mobile species \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\

@@

If our results were an example of sensory drive, we would have expected to see consistent differences between habitat-populations across islands (a pattern that would have been a compelling indicator of adaptation, \citealt{Losos2011}). This is because environmental conditions that may be relevant to color signal detectability such as light, temperature, moisture and vegetation, are consistent within the three main and clearly distinct habitat-types found across the sampled islands, i.e. beach scrub, primary coppice and mangroves \citep{Howard1950, Schoener1968}. Moreover, the patterns of divergence expected under a sensory drive scenario should be consistent with increased detectability given the local light conditions, such as the high contrasts with background vegetation found in the UV-range by \citet{Leal2002} and \citet{Leal2004}.\\

Instead, we found differences in the way dewlap color differs between habitats across islands. While short-wavelengths (UV reflectance) were often involved in color differences, they were not involved on all islands where significant differences were detected. On some islands, other or additional variables differed, such as brightness, red reflectance or the reflectance at the ends of the spectrum visible to \textit{Anolis} lizards (UV and red, \citealt{Lazareva2012}) relative to intermediate wavelengths (blue-to-yellow). Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than beach scrub lizards on the former, but a lower UV-reflectance on the latter. Under a sensory-drive scenario, we would have expected the same variables to be consistently divergent between habitats, or at least in a consistent direction.\\

Consistent patterns across islands would have been a good clue for a sensory-drive explanation, but in particular consistent differences between habitats that are most different in their local conditions regarding the ecological function of the dewlap, such as ambient light. For example, if ambient light is an important factor shaping dewlap coloration, we would expect mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to harbor more similar dewlaps, and to differ significantly from lizards from the coppice habitat, where irradiance is low. Overall, the observed heterogeneity of divergence patterns across islands provides no support to a sensory-drive explanation.\\

Phenotypic plasticity could be another cause for dewlap color variation between habitats, where different environmental conditions produce different phenotypes in different habitats. The yellow, orange and red colors in anoline dewlaps are produced by pterins and carotenoids \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}. Animals can synthesize pterins from nucleotides, but lack the ability to synthesize carotenoids \citep{Goodwin1984, Hill2002, Hill2006}. Different food qualities across sites within islands could therefore potentially cause detectable differences in coloration. Alternatively, more subtle effects on dewlap color could arise from developmental plasticity and depend, e.g., on differences in egg-rearing conditions. However, more data are needed to test these hypotheses, and although \citet{Cook2013} have shown plastic responses of dewlap color in response to parasites, we find it unlikely to account for the widespread habitat differences we found. Besides, studies testing the effect of carotenoid deprivation \citep{Steffen2010, Ng2013} and heritability \citep{Cox2017} of dewlap coloration in \textit{A. sagrei} and another species with a carotenoid-based dewlap, \textit{A. distichus}, found little support for phenotypic and developmental plasticity in dewlap coloration.\\

Genetic drift is another process that can account for differences in phenotype between localities, especially in small populations. One way this could proceed is through isolation-by-distance, where more distant populations accumulate more differences through time because of the reduced effect of gene flow at larger geographical scales relative to the dispersal range of the species \citep{Rousset2004}. Here, we only found a significant correlation between phenotypic and geographical distances on Eleuthera to support this scenario. On all the other islands, in contrast, populations from closer sites were not phenotypically more similar, which argues against isolation-by-distance. That said, there were often few sampling sites per island in our study, whose locations were not uniformly chosen within the islands, and so the true extent of isolation-by-distance may be difficult to test. Other forms of drift may be at play than isolation-by-distance, but nevertheless, we did find significant differences in color phenotype at relatively small spatial scales, sometimes in neighboring habitats, on islands where gene flow is probably highly pervasive, as suggested by high rates of encounter between males and females \citep{Kamath2018}, making the divergence of habitat-populations by drift in relative genetic isolation an unlikely scenario.\\

A number of alternative explanations remain. First, there could be sexual selection for different dewlap colors in different locations. Indeed, although the sensory-drive hypothesis may include a sexual selection aspect, e.g., if the optimal male phenotype in a given habitat is a function of female perception, sexual preferences may also be arbitrary and independent of the habitat \citep{Andersson1994}, and so could differ across habitats and islands. Although one previous study has found no link between dewlap coloration and body size dimorphism in \textit{A. sagrei} (a proxy for the intensity of sexual selection) in an among-island comparison \citep{Baeckens2018}, within-island data are lacking to test the plausibility of this scenario.\\

Alternatively, selective pressures may be different in similar habitats from one island to another, because of other environmental variables not accounted for by the habitat-type classification we used \citep{Howard1950, Schoener1968}. The islands we sampled exhibit variation in some climatic variables, but also in densities of predators and anole congeners, which have all been shown to correlate with variation in \textit{A. sagrei} mean dewlap coloration among islands \citep{Vanhooydonck2009, Baeckens2018}.\\

Finally, different island-populations could also respond differently to similar selective pressures, resulting in various between-habitat divergence patterns across islands. Several factors could account for this. For example, the founder populations of each island, which colonized the islands independently (van de Schoot, unpublished thesis; \citealt{Driessens2017, Reynolds2020}), could have exhibited different dewlap colors at the time of colonization, as may be suggested by the larger differences we observed between than within islands. In turn, different initial phenotypes could have led to different ways in which populations would have diverged between habitats. Moreover, the different founding populations may have also consisted in different subsets of the standing genetic variation of their Cuban ancestor due to potential bottlenecks \citep{Reynolds2020}, which may have constrained the way they would later respond to the local selective pressures of their respective islands.\\

Altogether, our results show that dewlap color of \textit{A. sagrei} commonly varies between habitat types, even in close geographical proximity, within islands of the West Indies, and that coloration differs in different ways from one island to another. We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to thoroughly test these hypotheses. Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.

>>>>

Detectable differences in dewlap color between populations are surprising, as habitats were often in close geographical proximity to each other (as close as a few hundred meters on Bimini and most of the time within ten kilometers). Indeed, given that (1) the populations were continuously distributed between the habitats, (2) populations from different habitats were not monophyletic with respect to mitochondrial haplotypes (\citealt{vandeSchoot2016} unpublished thesis), and (3) \textit{A. sagrei} has been shown to be a highly mobile species \citep{Kamath2018}, we would have expected more homogeneous distributions of color phenotypes within islands due to extensive gene flow, with fewer differences between populations, especially those in close proximity.\\

@@

If our results were an example of sensory drive, we would have expected to see consistent differences between populations from different habitats across islands, given the apparent environmental consistency each of the three habitat types across the islands we sampled. In particular, we would have expected divergence in line with increased detectability given local light conditions, such as the high contrasts with background vegetation found in the UV range in \citet{Leal2002} and \citet{Leal2004}. We might also have expected mangrove and beach scrub lizards, both inhabiting areas with high light penetration, to harbor more similar dewlaps, and to differ significantly form lizards from the coppice habitat, where irradiance is low. Instead, we found inconsistencies in the way dewlap color differed between habitats across islands. While short-wavelengths (UV reflectance) were often involved in color differences, they were not involved on all islands where significant differences were detected. On some islands, other or additional variables differed, such as brightness, red reflectance or the reflectance at the ends of the spectrum visible to \textit{Anolis} lizards (UV and red, \citealt{Lazareva2012}) relative to intermediate wavelengths (blue-to-yellow). Similar portions of the spectrum were sometimes involved in opposite directions on different islands, such as on Abaco and Cayman Brac, where mangrove lizards had a higher UV-reflectance than beach scrub lizards on the former, but a lower UV-reflectance on the latter. Overall, the observed heterogeneity of divergence patterns across islands provides no support to a sensory-drive explanation.\\

% Besides, pterins (the other orange pigment in dewlaps) can be synthesized \citep{Ortiz1962, Ortiz1962a, Ortiz1963, Ortiz1966, Macedonia2000, Steffen2007, Steffen2009}

It is presently not known if the reported differences in coloration have a genetic basis. Yet, we find it unlikely that these differences arose through phenotypic plasticity, as although the carotenoids that partly make up the red and orange colors of anole dewlaps must be found in the diet \citep{Goodwin1984, Hill2002, Hill2006}, studies testing the effect of carotenoid deprivation \citep{Steffen2010, Ng2013} and heritability \citep{Cox2017} of dewlap coloration in \textit{A. sagrei} and \textit{A. distichus} (another species with a carotenoid-based dewlap), found little support for phenotypic and developmental plasticity in dewlap coloration. One exception is a study demonstrating that lizards heavily parasitized by skin mites had duller dewlaps \citep{Cook2013}, but we found no sign of that in our study.\\

We found no evidence for a role of genetic drift in explaining the observed patterns either. First, \textit{A. sagrei} was distributed across the islands continuously, usually at relatively high population densities, and not in small and isolated populations where drift might be expected to have a strong effect. Second, we found no evidence of isolation-by-distance except on Abaco, which was sampled at the largest geographical scale, with sites nearly a hundred kilometers apart from each other. Hence, while isolation-by-distance may explain long-range differences on this island, most of the differences among habitats across islands, which were often in close proximity (less than 10km), are unlikely to be explained by genetic drift.\\

In this study, we found larger differences among than within islands, a pattern already reported and linked to climatic conditions \citep{Vanhooydonck2009} and to densities of predators and of anole congeners \citep{Baeckens2018}. Differences among habitats within islands, however, are still difficult to account for. Remaining hypotheses may include, for example, runaway sexual selection arbitrarily operating in different directions across islands \citep{Andersson1994}, but no evidence so far suggests that dewlap is a target of mate choice in anoles \citep{Tokarz2002, Tokarz2005, Nicholson2007}. Another hypothesis is that the different genetic constitutions of different islands, perhaps resulting from founder effects (the islands have been colonized independently from Cuba, \citealt{vandeSchoot2016} unpubl.; \citealt{Driessens2017, Reynolds2020}), may have predisposed populations to adapt differently to similarly selective circumstances. Either way, new data would be needed to test these hypotheses.\\

Altogether, our results show that dewlap color of \textit{A. sagrei} commonly varies between habitat types, even in close geographical proximity, within islands of the West Indies. Besides, coloration differs in different ways across similar habitats from one island to another. We discussed several non-mutually exclusive mechanisms that could explain these observations, but more data are needed to investigate these scenarios. Nevertheless, heterogeneous patterns of divergence across islands do not support an adaptive sensory-drive scenario, and we propose that within-island dewlap color variation may be underlain by a more subtle mosaic of factors.

### 2021-07-26 ms/sections/introduction.tex
v3.2

Here, we analyzed the color characteristics of \textit{A. sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands. These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this promiscuous species \citep{Kamath2018}. These islands all share three characteristic native West Indian small-island habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled. These habitats contrast in environmental parameters including vegetation community, light irradiance, humidity, and temperature \citep{Howard1950, Schoener1968}. The Cayman Islands and the Bahamas have been colonized independently by \textit{A. sagrei} from Cuba (\citealt{vandeSchoot2016} unpublished thesis; \citealt{Reynolds2020}), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\

Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km. Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A. sagrei}. We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest. Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats. Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Losos2011}.

>>>>

Here, we analyzed the color characteristics of \textit{A. sagrei} dewlaps within nine islands in the Bahamas and Cayman Islands. These island systems presently, if not historically, comprise relatively small islands, with no major geographic barriers within islands limiting dispersal for this species \citep{Kamath2018}. These islands all share three characteristic native West Indian small-island habitat-types -- beach scrub bush, closed-canopy primary coppice forest, and mangrove forest -- that are often spatially intermingled. These habitats contrast in environmental parameters including vegetation community, light irradiance, humidity, and temperature \citep{Howard1950, Schoener1968}. The Cayman Islands and the Bahamas have been colonized independently by \textit{A. sagrei} from Cuba (\citealt{vandeSchoot2016} unpublished thesis; \citealt{Reynolds2020}), such that these archipelagos constitute an ideal suite of natural replicates to explore within-island dewlap diversity across multiple islands.\\

Our sampling design included sites in close proximity; the median distance between two sites within an island was $11.2$km. Combining reflectance spectrometry and supervised machine learning, we tested for divergence in dewlap phenotype between habitats within islands and between islands across part of the range of \textit{A. sagrei}. We predicted that if light conditions in the environment indeed drive color evolution, dewlaps should be most similar between beach scrub and mangrove forest, which both have high levels of light irradiance, compared to the darker, closed-canopy coppice forest. Similar, if detectability is maximized given the local conditions, we expected darker and more contrasting dewlaps in high irradiance habitats. Finally, if habitat characteristics are strong determinants of dewlap color variation, similar patterns should be observed across multiple islands \citep{Harvey1991, Losos2011}.

### 2021-07-26 ms/sections/methods2.tex
v3.2

We sampled 466 male \textit{A. sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, and Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A). These islands were chosen to span the breadth of the West Indian range of \textit{A. sagrei}, because they have highly similar habitat types,  and because the \textit{A. sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e. relatively evolutionarily independent, \citealt{Reynolds2020}). Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by its dominant vegetation type --- xeric beach scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration). Sample sizes are given in Table \ref{tab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $\sim 11$km (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km. Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

@@

We tested for detectable differences in dewlap coloration between populations from different habitats across islands by following an analytic pipeline in several steps. First, we used multivariate analyses of variance to assess the relative contributions of islands, habitats and habitat-by-island interactions on the partitioning of variation in color space. Second, and provided that habitat-by-island interactions were found, we investigated habitat-differences in dewlap color for each island separately using machine learning classification. Third, for each island where multivariate differences were detected using our machine learning pipeline, we used univariate analyses of variance to decompose the signal among the different dimensions of color space. Fourth, for each significant between-habitat variation found in univariate analyses, we used post-hoc tests to determine which habitats were responsible for the differences. Last, to get insights into the spatial scale of phenotypic variation, for each significant contrast between two habitats detected along a given dimension on a given island, we performed multiple pairwise Wilcoxon tests to assess differences in dewlap coloration among the sites involved in that significant contrast, and recorded the geographical distance between sites that were found significant. In parallel, we tested a possible effect of isolation-by-distance, an alternative cause of phenotypic divergence between localities, based on diffusion approximation and dispersal distance, irrespective of habitat-types. We did so using a permutation test to assess the significance of the correlation between geographical distances and phenotypic distances among sites within each island.\\

@@

For each island where significant differences in dewlap coloration were detected between habitats, we used multiple univariate analyses of variance (ANOVA) to identify possible principal components underlying the observed differences. We constructed our ANOVA models in two steps, as per \citet{Zuur2009}. In a first step, we accounted for heterogeneity of variances across groups by systematically comparing the goodness-of-fit of an ANOVA model estimated with ordinary least squares (OLS) with that of a model estimated with generalized least squares (GLS), which allowed one estimate of residual variance per habitat (using the R package nlme, \citealt{Pinheiro2000, Pinheiro2020}). Both models were fitted with restricted maximum likelihood (REML). Goodness-of-fit was estimated using Akaike's Information Criterion corrected for small sample sizes (AICc, R package MuMIn, \citealt{Barton2019}), and the estimation method yielding the lowest AICc was retained. In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat-type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking the habitat-term.\\

@@

We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their Euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{tab:sites}), regardless of habitat-type. Because often only a few sites were sampled per island, we could not get meaningful results from tests that use sites as units of observation, such as Moran's I test \citep{Gittleman1990}. Instead, we designed a permutation test where we randomly reshuffled individual lizards across sites within islands 1,000 times each, and systematically recalculated Pearson's correlation coefficient between geographic distances (computed as geodesic distances in the R package geosphere; \citealt{Hijmans2019}) and phenotypic distances. We used the resulting null distributions of correlation coefficients to assess the significance of the observed spatial autocorrelation for each island.

@@

In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands. We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat-types, for each pair of habitats and each variable where significant differences were detected with our analyses of variance. The P-values were adjusted using a Benjamini-Hochberg correction for multiple testing \citep{Benjamini1995}.

>>>>

We sampled 466 male \textit{A. sagrei} from seven islands in the Bahamas Archipelago -- Abaco, North Andros, South Andros, South Bimini, Eleuthera, Long Island, and Ragged Island -- and two in the Cayman Islands -- Cayman Brac and Little Cayman (Figure \ref{fig:maps}A). These islands were chosen to span the breadth of the West Indian range of \textit{A. sagrei}, because they have highly similar habitat types,  and because the \textit{A. sagrei} on each island group are derived from ancient and distinct colonization events from Cuba (i.e. relatively evolutionarily independent, \citealt{Reynolds2020}). Three habitats were sampled on each island based on characterizations by \citet{Howard1950} and \citet{Schoener1968}. Each habitat is clearly distinguishable by its dominant vegetation type --- xeric beach scrub (open, relatively dry habitat consisting of low vegetation or isolated trees), primary coppice forest (closed-canopy forest) and mangrove forest (wet coastal habitat with trees growing in brackish water and high light penetration, although lizards were sampled in dry soil areas). Sample sizes are given in Table \ref{tab:counts}. Our sampling design enabled us to test for differences between habitats at a coarse and fine geographical scale. The median distance between two localities within an island was $\sim 11$km (Figure \ref{fig:maps}B), and $80.3$\% of all pairwise distances within islands were less than $50$km. Additionally, there are no major barriers to dispersal (such as mountains or grassland) on any of the islands that we sampled.

@@

We tested for detectable differences in dewlap coloration between populations from different habitats across islands by following an analytic pipeline in several steps. First, we used multivariate analyses of variance to assess the relative contributions of islands, habitats and habitat-by-island interactions on the partitioning of variation in color space. Second, and provided that habitat-by-island interactions were found, we investigated habitat-differences in dewlap color for each island separately using machine learning classification. Third, for each island where multivariate differences were detected using our machine learning pipeline, we used univariate analyses of variance to decompose the signal among the different dimensions of color space. Fourth, for each significant between-habitat variation found in univariate analyses, we used post-hoc tests to determine which habitats were responsible for the differences. Last, to get insights into the spatial scale of phenotypic variation, for each significant contrast between two habitats detected along a given dimension on a given island, we performed multiple pairwise Wilcoxon tests to assess differences in dewlap coloration among the sites involved in that significant contrast, and recorded the geographical distance between sites that were found significant. In parallel, we tested a possible effect of isolation-by-distance, an alternative cause of phenotypic divergence between localities, based on diffusion approximation and dispersal distance, irrespective of habitat types. We did so using a permutation test to assess the significance of the correlation between geographical distances and phenotypic distances among sites within each island.\\

@@

For each island where significant differences in dewlap coloration were detected between habitats, we used multiple univariate analyses of variance (ANOVA) to identify possible principal components underlying the observed differences. We constructed our ANOVA models in two steps, as per \citet{Zuur2009}. In a first step, we accounted for heterogeneity of variances across groups by systematically comparing the goodness-of-fit of an ANOVA model estimated with ordinary least squares (OLS) with that of a model estimated with generalized least squares (GLS), which allowed one estimate of residual variance per habitat (using the R package nlme, \citealt{Pinheiro2000, Pinheiro2020}). Both models were fitted with restricted maximum likelihood (REML). Goodness-of-fit was estimated using Akaike's Information Criterion corrected for small sample sizes (AICc, R package MuMIn, \citealt{Barton2019}), and the estimation method yielding the lowest AICc was retained. In a second step, we re-fitted the retained model with maximum likelihood (ML) to test for the effect of habitat type using likelihood ratio tests (LRT) between a model including a habitat-term and a null model lacking the habitat-term.\\

@@

We tested for within-island spatial autocorrelation between the geographical distances among sampling sites and their Euclidean distances in multivariate color space (mean PC1 to PC4 per site, Table \ref{tab:sites}), regardless of habitat type. For this, we performed Mantel's test (\citealt{Legendre2012}, R package vegan; \citealt{Oksanen2019}) on each island, using 999 permutations and geographical distances computed as geodesic distances from latitude and longitude data (R package geosphere, \citealt{Hijmans2019}).\\

@@

In this study, we were interested in the minimum spatial scale at which significant differences between habitats could be detected within islands. We performed multiple pairwise nonparametric Wilcoxon-Mann-Whitney tests \citep{Hollander2013} to compare dewlap coloration between sites with different habitat types, for each pair of habitats and each variable where significant differences were detected with our analyses of variance. The P-values were adjusted using a Benjamini-Hochberg correction for multiple testing \citep{Benjamini1995}.
